---
title: "Venn Diagram"
output: html_notebook
---

```{r}
library(VennDiagram)
install.packages("extrafont")
library(extrafont)
install.packages("Cairo")
#font_import()
```
Two-sided Venn
```{r}
library("VennDiagram")

grid.newpage()  
two.way.venn <- draw.pairwise.venn(
  area1 = 1575,
  area2 = 1227,
  cross.area = 613,
  col = c("79380", "800"),
  fill = c("#fc03e8", "#ED1C24"),
  fontfamily = "Arial",
  category.cex = 1,  ##size
  category.fontface = "regular",  ##style
  euler.d.fontface = "regular",  ##style for intersection label
  euler.d.cex = 1.5,  ##size for intersection
  lty = "blank",
                 cex = 0.8,       # Adjust the text size (e.g., 1.5 for 50% larger)
                 cat.cex = 1)   # Adjust the category label text size, if needed)
png(filename = "chol-vs-palm_JF570_venn_14dec24.png", width = 2000, height = 1400, res = 1400)
grid.draw(two.way.venn)
dev.off()

                  
```
area1 <- 485
area2 <- 297
area3 <- 221
n12 <- 126
n13 <- 165
n23 <- 82
n123 <- 71

```{r}
library(VennDiagram)

grid.newpage()
venn.plot <- draw.triple.venn(area1 = 485,
                 area2 = 221,
                 area3 = 484,
                 n12 = 165,       
                 n23 = 122,       
                 n13 = 185,
                 n123 = 106,
                 category = c("","",""),
                 fill = c("#be4a80", "#67a5c6", "black"),
                 lty = "blank",
                 cex = 0,       # Adjust the text size (e.g., 1.5 for 50% larger)
                 cat.cex = 1)   # Adjust the category label text size, if needed)
png(filename = "3-way-venn_HDL_cholJF570-heps_7Aug24.png", width = 2000, height = 1400, res = 1400)
grid.draw(venn.plot)
dev.off()

```


Quadrupule venn diagram
Making the venn diagram. I have already done all the "heavy lifting" in excel using the vlookup function.
```{r}
grid.newpage()
venn.plot <- draw.quad.venn(area1 = 587,      #AP2-123b
                 area2 = 558,    #LMK38
                 area3 = 503,    #NBII-165
                 area4 = 754,    #trans-sterol (PRC)
                 n12 = 154,       
                 n23 = 347,       
                 n13 = 130,
                 n14 = 236,
                 n24 = 379,
                 n34 = 344,
                 n123 = 100,
                 n124 = 137,
                 n234 = 262,
                 n134 = 122,
                 n1234 = 93,
                 category = c("chol-JF570","LKM38","NBII-165","trans-sterol"),
                 fill = c("#fc03e8", "green", "orange", "blue"),
                 lty = "blank",
                 cex = 2,       # Adjust the text size (e.g., 1.5 for 50% larger)
                 cat.cex = 1)   # Adjust the category label text size, if needed
grid.draw(venn.plot)
dev.off()
```


```{r}
#SWITCHING 2<->4
grid.newpage()
venn.plot <- draw.quad.venn(area1 = 1575,      #EB-B-012 chol
                 area2 = 1443,    #trans-sterol (PRC)
                 area3 = 452,    #NBII-165
                 area4 = 408,    #LKM38
                 n12 = 422,       
                 n23 = 283,       
                 n13 = 123,
                 n14 = 101,
                 n24 = 243,
                 n34 = 213,
                 n123 = 109,
                 n124 = 90,
                 n234 = 146,
                 n134 = 55,
                 n1234 = 52,
                 category = c("","","",""),
                 fill = c("#fc03e8", "dark green", "orange", "blue"),
                 lty = "blank",
                 cex = 0.6,       # Adjust the text size (e.g., 1.5 for 50% larger)
                 cat.cex = 1)   # Adjust the category label text size, if needed

png(filename = "4-way-venn_chol-diaz_JF570_9jan25.png", width = 2000, height = 1400, res = 1400)
grid.draw(venn.plot)
dev.off()

```

```{r}
# Load required libraries
library(VennDiagram)
library(Cairo)

# Create a file connection to the EPS file
output_file <- "venn_diagram.eps"
Cairo(file = output_file, width = 7, height = 7, type = "eps", units = "in", dpi = 300)

# Draw the Venn diagram with the "Times" font family
venn.plot <- draw.quad.venn(area1 = 765,
                 area2 = 754,
                 area3 = 503,
                 area4 = 558,
                 n12 = 273,       
                 n23 = 379,       
                 n13 = 142,
                 n14 = 163,
                 n24 = 379,
                 n34 = 347,
                 n123 = 132,
                 n124 = 147,
                 n234 = 262,
                 n134 = 104,
                 n1234 = 102,
                 category = c("chol-JF570","trans-sterol","NBII-165","LKM38"),
                 fill = c("magenta", "dark green", "orange", "blue"),
                 lty = "blank",
                 cat.fontfamily = "Times")  # Change the font family here


# Close the file connection
dev.off()

```







```{r}
# Writing to file
tiff(
    filename = tempfile(
        pattern = 'Quad_Venn_diagram',
        fileext = '.tiff'
        ),
    compression = "lzw"
    );
grid.draw(venn.plot);
dev.off()
```



```{r}
png(filename = "venn_plot.png", width = 1200, height = 1200, res = 1200)
grid.draw(venn.plot)
dev.off()

```




category    A vector (length 4) of strings giving the category names of the sets
lwd     A vector (length 4) of numbers giving the line width of the circles’ circumferences
lty     A vector (length 4) giving the dash pattern of the circles’ circumferences
col     A vector (length 4) giving the colours of the circles’ circumferences
fill    A vector (length 4) giving the colours of the circles’ areas
alpha     A vector (length 4) giving the alpha transparency of the circles’ areas
label.col     A vector (length 15) giving the colours of the areas’ labels
cex     A vector (length 15) giving the size of the areas’ labels
fontface    A vector (length 15) giving the fontface of the areas’ labels
fontfamily    A vector (length 15) giving the fontfamily of the areas’ labels
cat.pos     A vector (length 4) giving the positions (in degrees) of the category names along
the circles, with 0 (default) at 12 o’clock
cat.dist    A vector (length 4) giving the distances (in npc units) of the category names
from the edges of the circles (can be negative)
cat.cex     A vector (length 4) giving the size of the category names
cat.col     A vector (length 4) giving the colours of the category names
cat.fontface    A vector (length 4) giving the fontface of the category names
cat.fontfamily    A vector (length 4) giving the fontfamily of the category names
cat.just    List of 4 vectors of length 2 indicating horizontal and vertical justification of
each category name
rotation.degree   Number of degrees to rotate the entire diagram
rotation.centre   A vector (length 2) indicating (x,y) of the rotation centre
ind Boolean indicating whether the function is to automatically draw the diagram
before returning the gList object or not
cex.prop    A function or string used to rescale areas
print.mode    Can be either ’raw’ or ’percent’. This is the format that the numbers will be
printed in. Can pass in a vector with the second element being printed under the
first
sigdigs     If one of the elements in print.mode is ’percent’, then this is how many significant digits will be kept
direct.area     If this is equal to true, then the vector passed into area.vector will be directly
assigned to the areas of the corresponding regions. Only use this if you know
which positions in the vector correspond to which regions in the diagram
area.vector     An argument to be used when direct.area is true. These are the areas of the
corresponding regions in the Venn Diagram
